---
title: "Lab 0: Getting Started with Datahub and R"
subtitle: "POLI 172 - Advanced Social Data Analytics"
date: "2026-09-24"
date-format: "MMMM D, YYYY"
format: pdf
pdf-engine: xelatex
keep-tex: false
fontsize: "11pt"
geometry: "margin=0.85in"
execute:
  echo: true
  warning: false
  message: false
  error: false
  cache: false
  freeze: false
---

<!-- Instructor authoring: generated question version from lab00-solutions.qmd; edit the answer master and regenerate. Learners may edit their own question copy. -->


**Name:** [Your name]\
**PID:** [Your PID]

Work **individually** during class. 
Complete any remaining work and submit your **PDF to Gradescope by
Friday, September 25, at 11:59 p.m. (Pacific time)**.

Practice basic R commands from DSS Chapter 1 (pp. 8-25) and render this
Quarto (`.qmd`) document to PDF.

# Part 1: Get started with Quarto

## Exercise 1.1: Get started

Open `lab00.qmd` in RStudio's **Source** editor on
[Datahub](https://datahub.ucsd.edu/).
Enter your name and PID above, save, and
click **Render**. Find the following code and its result in the PDF.

```{r}
#| label: exercise-1-1-first-render
#| eval: true
1 + 3
```

If setup or rendering fails, ask the instructor or PLA for help during class.

**Where to write code and answers**

A QMD file combines **R code chunks** and **Markdown text**. In Source, find
the chunk around `1 + 3` above: it opens with three backticks and an r in
braces, and closes with three backticks.

- **Code:** Inside each exercise chunk, replace `# YOUR CODE HERE` with R
  commands. Complete and enable earlier chunks first. Keep backticks and
  `#| label:` lines unchanged.
- **Written answers:** Outside chunks, replace `[Your answer here]` with
  ordinary sentences. Markdown adds formatting: `**bold**` makes text bold;
  `## Heading` creates a heading.

Leave the document settings at the top unchanged. **Render** runs enabled
chunks from top to bottom and combines your text, code, and output in a PDF.

**Help:** [Quarto in RStudio](https://quarto.org/docs/get-started/hello/rstudio.html#authoring)
and [Markdown basics](https://quarto.org/docs/authoring/markdown-basics.html).


# Part 2: Use R as a calculator

R uses `+`, `-`, `*`, and `/` for arithmetic. Parentheses control which
calculation happens first. Inside an R chunk, `#` starts a comment for R.
Lines beginning with `#|` also supply settings for rendering.

**Keep chunk labels:** `#| label: exercise-2-1-calculator` identifies
Exercise 2.1 in render progress and R-code error messages. **Do not delete
or rename label lines**; they help us locate problems when rendering fails.

## Exercise 2.1: Calculate and compare

Use R to calculate both expressions below, on separate lines:

- `(4 - 1 + 3) * (2 / 3)`
- `4 - 1 + 3 * 2 / 3`

```{r}
#| label: exercise-2-1-calculator
#| echo: true
#| eval: true
# YOUR CODE HERE
```

**Question:** Why do the expressions give different results? Use one sentence.

[Your answer here]

## Exercise 2.2: Create a vector and find its mean

`<-` stores a value in an object. `c()` combines values into a vector, and
`mean()` calculates their average. The supplied line stores three values
in a vector named `x`.

Run that line, then add code to **print `x`** and **calculate its mean**.
To print an object, write its name on a line by itself; to find its mean,
put its name inside `mean()`.

```{r}
#| label: exercise-2-2-vector-mean
#| echo: true
#| eval: true
x <- c(10, 8, 12)
# YOUR CODE HERE
```

**Question:** What is the mean for these three values?

[Your answer here]


# Part 3: Load election data

`elections.csv` contains U.S. presidential election results. Each row records
the votes in one county in one election year. Keep the CSV and QMD in the
same folder.

| Variable | Meaning |
|---|---|
| `year` | Election year |
| `state`, `county` | State and county names |
| `rep` | Votes for the Republican candidate |
| `dem` | Votes for the Democratic candidate |
| `other` | Votes for other candidates |

## Exercise 3.1: Read and preview the data

Load `elections.csv` using `read.csv()` and store it in an object named
`elections`. Then use `head()` to print its first six rows.

**Hint:** `object <- read.csv("filename.csv")` loads a file;
`head(object)` previews it. Keep the CSV and QMD in the same folder.

```{r}
#| label: exercise-3-1-load-preview
#| echo: true
#| eval: true
# YOUR CODE HERE
```

## Exercise 3.2: Compute and interpret a mean

Use `mean()` to find the average number of Democratic votes (`dem`) across
**all rows** of `elections`. The expression `elections$dem` selects that column.
Complete and enable Exercise 3.1 first.

```{r}
#| label: exercise-3-2-data-mean
#| echo: true
#| eval: true
# YOUR CODE HERE
```

**Question:** In one sentence, report the average number of Democratic votes
per county per election in this dataset.

[Your answer here]

# Part 4: Render and submit

1. Save your work. 
2. Click **Render**. Open the PDF and check your name, code, output,
   and written answers. 
3. In **Files**, select `lab00.pdf`, then **More > Export** to download it. 
   Submit that PDF to Gradescope by **Friday at 11:59 p.m.** Keep your `.qmd` for future edits.
